### 2018-10-31 Sections/physical_architectures_for_quantum_computing.tex
Boson-sampling permanents

The permanent of a square matrix is defined as\index{Permanents},
\begin{align}
\mathrm{Per}(A) = \sum_{\sigma\in S_n} \prod_{i=1}^n A_{i,\sigma_i},
\end{align}
which sums over $n!$ terms (super-exponential). Note the similarity with the definition for the matrix determinant\index{Determinants}, defines identically, but with the addition of an alternating $\pm$-sign in the terms. Despite this similarity, determinants reside in \textbf{P}, with an efficient classical algorithm. For this reason, Fermionic sampling\index{Fermionic sampling} yields an easy computational problem, since Fermionic sampling differs only in replacing the permanent with the determinant. The best-known classical algorithm for evaluating has exponential runtime,
\begin{align}
	O(2^n n^2).
\end{align}

To see how matrix permanents naturally arise in this setting, it is easiest to explain by example. In Fig.~\ref{fig:BS_2_comb} we illustrate a simple interferometer, fed with two photons. We wish to calculate the output amplitude of measuring a photon in each of the modes 2 and 3, given photons input at modes 1 and 2. To evaluate this amplitude we simply need to add up the amplitudes of all possible paths yielding the desired outcome. In this simple example this sum-of-paths\index{Sum-of-paths} is given by,
\begin{align} \label{eq:BS_2_ph_comb}
\gamma_{\{2,3\}} &= \underbrace{U_{1,2}U_{2,3}}_{\mathrm{don't\ swap}} + \underbrace{U_{1,3}U_{2,2}}_{\mathrm{swap}} \nonumber \\
&= \mathrm{Per} \left[ {\begin{array}{cc}
   U_{1,2} & U_{2,2} \\
   U_{1,3} & U_{2,3} \\
  \end{array} } \right],
\end{align}
from which it is immediately clear that the amplitude is given by the permanent of the \mbox{$2\times 2$} matrix obtained from taking the columns (rows) of $\hat{U}$ where a photon is present at the respective input (output) mode.

@@

In Fig.~\ref{fig:BS_3_comb} we present to next most sophisticated example of an interferometer fed by 3 photons, for which the sum-of-paths\index{Sum-of-paths} is given by,
\begin{align} \label{eq:BS_3_ph_comb}
\gamma_{\{1,2,3\}} &= U_{1,1}U_{2,2}U_{3,3} + U_{1,1}U_{3,2}U_{2,3} \nonumber \\
&+ U_{2,1}U_{1,2}U_{3,3} + U_{2,1}U_{3,2}U_{1,3} \nonumber \\
&+ U_{3,1}U_{1,2}U_{2,3} + U_{3,1}U_{2,2}U_{1,3}
\nonumber \\
&= \mathrm{Per} \left[ {\begin{array}{ccc}
   U_{1,1} & U_{2,1} & U_{3,1} \\
   U_{1,2} & U_{2,2} & U_{3,2} \\
   U_{1,3} & U_{2,3} & U_{3,3} \\
  \end{array} } \right],
\end{align}
and it is now clear upon inspection that the amplitude is given by a \mbox{$3\times 3$} matrix permanent.

>>>>

The permanent of a square matrix is defined as\index{Permanents},
\begin{align}\label{eq:permanent}
\mathrm{Per}(A) = \sum_{\sigma\in S_n} \prod_{i=1}^n A_{i,\sigma_i},
\end{align}
which sums over $n!$ terms (super-exponential), where $S_n$ is the symmetric group, the group of permutations on $n$ elements, of which there are $n!$. Note the similarity with the definition for the matrix determinant\index{Determinants}, defines identically, but with the addition of an alternating $\pm$-sign in the terms. Despite this similarity, determinants reside in \textbf{P}, with an efficient classical algorithm. For this reason, Fermionic sampling\index{Fermionic sampling} yields an easy computational problem, since Fermionic sampling differs only in replacing the permanent with the determinant. The best-known classical algorithm for evaluating has exponential runtime,
\begin{align}
	O(2^n n^2).
\end{align}

To see how matrix permanents naturally arise in this setting, it is easiest to explain by example. In Fig.~\ref{fig:BS_2_comb} we illustrate a simple interferometer, fed with two photons. We wish to calculate the output amplitude of measuring a photon in each of the modes 2 and 3, given photons input at modes 1 and 2. To evaluate this amplitude we simply need to add up the amplitudes of all possible paths yielding the desired outcome. In this simple example this sum-of-paths\index{Sum-of-paths} is given by,
\begin{align} \label{eq:BS_2_ph_comb}
\gamma_{\{2,3\}} &= \underbrace{U_{1,2}U_{2,3}}_{\mathrm{don't\ swap}} + \underbrace{U_{1,3}U_{2,2}}_{\mathrm{swap}} \nonumber \\
&= \mathrm{Per} \left[ {\begin{array}{cc}
   U_{1,2} & U_{2,2} \\
   U_{1,3} & U_{2,3} \\
  \end{array} } \right],
\end{align}
from which it is immediately clear that the amplitude is given by the sum of \mbox{$2!=2$} paths\footnote{The number of paths scales as $n!$ in general, which corresponds to the $n!$ order of the symmetric group, $S_n$, in the definition of the permanent from Eq.~(\ref{eq:permanent}).}, the permanent of the \mbox{$2\times 2$} matrix obtained from taking the columns (rows) of $\hat{U}$ where a photon is present at the respective input (output) mode.

@@

In Fig.~\ref{fig:BS_3_comb} we present to next most sophisticated example of an interferometer fed by 3 photons, for which the sum-of-paths\index{Sum-of-paths} has \mbox{$3!=6$} terms, given by,
\begin{align} \label{eq:BS_3_ph_comb}
\gamma_{\{1,2,3\}} &= U_{1,1}U_{2,2}U_{3,3} + U_{1,1}U_{3,2}U_{2,3} \nonumber \\
&+ U_{2,1}U_{1,2}U_{3,3} + U_{2,1}U_{3,2}U_{1,3} \nonumber \\
&+ U_{3,1}U_{1,2}U_{2,3} + U_{3,1}U_{2,2}U_{1,3}
\nonumber \\
&= \mathrm{Per} \left[ {\begin{array}{ccc}
   U_{1,1} & U_{2,1} & U_{3,1} \\
   U_{1,2} & U_{2,2} & U_{3,2} \\
   U_{1,3} & U_{2,3} & U_{3,3} \\
  \end{array} } \right],
\end{align}
and it is now clear upon inspection that the amplitude is given by a \mbox{$3\times 3$} matrix permanent.
